---
title: "Social Network Analytics : International Alliance Networks"
author: "-Michael Schweizer, Mingxin Yin, Nupur K. Neti, Xinrong Chen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(igraph)
library(sna)
library("xergm.common")
library("networkDynamic")
library("btergm")
```

# Data Description

We decided to use the dataset "alliances" from the xergm.common package. It contains the international defense alliance network among 164 countries, covering the years 1981–2000. This is an excerpt from a dataset that has been used in two published analyses. The full dataset (Cranmer, Desmarais and Menninga 2012; Cranmer, Desmarais and Kirlkand 2012) contains a large number of countries and a much longer time series.

The alliances dataset contains 5 main data structures:

### allyNet
is a list of network objects at 20 time points, 1981–2000, containing undirected defense alliance networks. A 1 indicates an alliance and 0 indicates no alliance. In addition to the yearly defense alliance network, it contains attributes for  military capabilities (CINC) and governing regime type ("polity"). A higher CINC (Composite Index of National Capability) score indicates a more capable military state.Polity Score captures te regime authority spectrum on a 21-point scale ranging from -10 (hereditary monarchy) to +10 (consolidated democracy).

### contigMat
A 164 x 164 binary matrix in which a 1 indicates that two countries share a border.

### lNet
A list of 20 matrices. Each element is the adjacency matrix from the previous year. This is used to model memory in the ties.

### LSP
A list of 20 matrices. Each element is a matrix recording the number of shared partners between countries in the alliance network from the previous year.

### warNet
A list of 20 matrices. Each element is a binary matrix that indicates whether two states were in a militarized interstate dispute in the respective year.

```{r}
#extracting data
data("alliances", package = "xergm.common")
```

# Interesting questions we wanted to answer

* Do geographic factors influence alliances and war?
* How do alliances change over time?
* What influences the formation and dissolution of alliances?
* Do communities of alliances exist? If so, are they geographically linked?
* Does a history of war influnce alliances in the future?
* Do state capabilities and regimes make a country more attractive as an alliance partner?

# Data preparation

We chose to look at the years 1985 and 1995 since these were turbulent years due to the dissolution of USSR and we thought we could find some interesting patterns in the alliance network between these years. 

```{r}
#extracting attribute data for 1985 and 1995
ally_85 <- allyNet[[5]]
ally_95 <- allyNet[[15]]
ally_85_cinc <- ally_85%v%"cinc"
ally_95_cinc <- ally_95%v%"cinc"
ally_85_polity <- ally_85%v%"polity"
ally_95_polity <- ally_95%v%"polity"
```

```{r}
#extracting relationship data for 1985 and 1995
LSP_85 <- as.matrix(LSP[[5]])
LSP_95 <- as.matrix(LSP[[15]])
war_85 <- as.matrix(warNet[[5]])
war_95 <- as.matrix (warNet[[15]])
ally_85 <- as.matrix(allyNet[[5]])
ally_95 <- as.matrix(allyNet[[15]])

```

```{r}
#creating attrs df
attrs85 <- data.frame(cinc = ally_85_cinc, ally = ally_85_polity)
attrs95 <- data.frame(cinc = ally_95_cinc, ally = ally_95_polity)
rownames(attrs85) <- rownames(war_85)
rownames(attrs95) <- rownames(war_95)
```

```{r}
head(LSP_85[,0:20])
head(war_85[,0:20])
head(ally_85[,0:20])
```

## Network Metrics

## War Dataset
Each element is a binary matrix that indicates whether two states were in a militarized interstate dispute in the respective year.
```{r}
g_WAR_85<- graph.adjacency(war_85, weighted = NULL, mode = 'undirected')
g_WAR_95<- graph.adjacency(war_95, weighted = NULL, mode = 'undirected')
```

#The number of vertices and the edge count (network size)
```{r}
vcount(g_WAR_85)
ecount(g_WAR_85)
vcount(g_WAR_95)
ecount(g_WAR_95)
```
164 countries are included in the dataset. There are 171 ongoing war in 1985 and 192 in 1995.

#Network density
Network density is the portion of the potential connections in a network that are actual connections.In this context it is the probability that two random countries are at war. 
```{r}
edge_density(g_WAR_85, loops=TRUE)
edge_density(g_WAR_95, loops=TRUE)
```
The number increased from 1985 to 1995, indicating the world tension increases a little bit.

#Degree distribution
```{r}
degree_WAR_85 <- igraph::degree(g_WAR_85)
nodeName <- V(g_WAR_85)$name
degree_dist_WAR_85 <- data.frame(nodeName, degree_WAR_85)
degree_dist_WAR_85 <- degree_dist_WAR_85[order(-degree_WAR_85),]
head(degree_dist_WAR_85,10)
degree_WAR_95 <- igraph::degree(g_WAR_95)
nodeName <- V(g_WAR_95)$name
degree_dist_WAR_95 <- data.frame(nodeName, degree_WAR_95)
degree_dist_WAR_95 <- degree_dist_WAR_95[order(-degree_WAR_95),]
head(degree_dist_WAR_95,10)
```
USA was involved in war with the largest number of countries (15) in 1985.
Iraq was involved in war with the largest number of countries (31) in 1995.

#Difference in degree distribution
```{r}
degree_dist_WAR <- merge(x = degree_dist_WAR_85, y = degree_dist_WAR_95, by ="nodeName", all.x = TRUE)
degree_dist_WAR$difference <- (degree_dist_WAR$degree_WAR_95 - degree_dist_WAR$degree_WAR_85)
head(degree_dist_WAR,10)
```
Iraq is the country with the biggest increase. (6 in 1985 to 31 in 1995)
Cuba is the country wuth the biggest decrease. (10 in 1985 and 2 in 1995)

#Plot
```{r}
set.seed(1234)
plot(g_WAR_85)
```

```{r}
set.seed(1234)
plot(g_WAR_95)
```

#Transitivity
Local clustering coefficient here represents the likelihood of countries who are in war with the same particular country also have a militarized conflict with each other.
```{r}
LocalCC_WAR_85<- transitivity(g_WAR_85, type="local")
LocalCC_WAR_85_df<- data.frame(nodeName, LocalCC_WAR_85)
LocalCC_WAR_85_df <- LocalCC_WAR_85_df[order(-LocalCC_WAR_85),]
head(LocalCC_WAR_85_df,10)
LocalCC_WAR_95<- transitivity(g_WAR_95, type="local")
LocalCC_WAR_95_df<- data.frame(nodeName, LocalCC_WAR_95)
LocalCC_WAR_95_df <- LocalCC_WAR_95_df[order(-LocalCC_WAR_95),]
head(LocalCC_WAR_95_df,10)
```

This value is 1 for Canada, Ecuador, Malaysia and Jordan have a high transitivity value of 1 in 1985. 
Similarly, Macedonia, Bosnia, Tunisia, Kuwait, Bahrain, United Arab Emirates and the Republic of Korea in 1995 have a transitivity of 1. 

That is, if they and another country are in war with the same particular country, they definitely have a militarized conflict with each other. We see groups of country having a high transitivity in the middle east in 1995. This tells us that maybe there was a highly clustered conflict in the area at the time. 

#Global clustering coefficient (CC)
global - ratio of triangles (direction disregarded) to connected triples.
```{r}
head(transitivity(g_WAR_85),10)
head(transitivity(g_WAR_95),10)
```

These values are low and tell us the war network was not very liekly to form clusters of wars. 
## Ally Dataset
Ally dataset contains undirected dense alliance networks.
```{r}
g_ALLY_85<- graph.adjacency(ally_85, weighted = NULL, mode = 'undirected')
g_ALLY_95<- graph.adjacency(ally_95, weighted = NULL, mode = 'undirected')
```

#The number of vertices and the edge count (network size)
```{r}
vcount(g_ALLY_85)
ecount(g_ALLY_85)
vcount(g_ALLY_95)
ecount(g_ALLY_95)
```
Number of countries in a defense alliance network is 623 in 1985 and 674 in 1995.

#Network density
```{r}
head(edge_density(g_ALLY_85, loops=TRUE),10)
head(edge_density(g_ALLY_95, loops=TRUE),10)
```
The probability that two random countries are allied with each other increase from 0.046 in 1985 to 0.049 in 1995

#Degree distribution
```{r}
degree_ALLY_85 <- igraph::degree(g_ALLY_85)
nodeName <- V(g_ALLY_85)$name
degree_dist_ALLY_85 <- data.frame(nodeName, degree_ALLY_85)
degree_dist_ALLY_85 <- degree_dist_ALLY_85[order(-degree_ALLY_85),]
head(degree_dist_ALLY_85,10)
degree_ALLY_95 <- igraph::degree(g_ALLY_95)
nodeName <- V(g_ALLY_95)$name
degree_dist_ALLY_95 <- data.frame(nodeName, degree_ALLY_95)
degree_dist_ALLY_95 <- degree_dist_ALLY_95[order(-degree_ALLY_95),]
head(degree_dist_ALLY_95,10)
```
USA has the most number of allies in both 1985 and 1995

#Difference in degree distribution
```{r}
degree_dist_ALLY <- merge(x = degree_dist_ALLY_85, y = degree_dist_ALLY_95, by ="nodeName", all.x = TRUE)
degree_dist_ALLY$difference <- (degree_dist_ALLY$degree_ALLY_95 - degree_dist_ALLY$degree_ALLY_85)
head(degree_dist_ALLY,10)
```
Canada is the country with the biggest increase. (13 in 1985 to 36 in 1995)
YPR is the country wuth the biggest decrease. (19 in 1985 to 0 in 1995)

#Plot
```{r}
set.seed(1234)
plot(g_ALLY_85)
```

```{r}
set.seed(1234)
plot(g_ALLY_95)
```

Notice that subgroup clustered together in the lower right corner of the 1985 plot disappears in 1995.

#Transitivity
Transitivity refers to the extent to which the relation that relates two nodes in a network that are connected by an edge is transitive.
```{r}
LocalCC_ALLY_85<- transitivity(g_ALLY_85, type="local")
LocalCC_ALLY_85_df<- data.frame(nodeName, LocalCC_ALLY_85)
LocalCC_ALLY_85_df <- LocalCC_ALLY_85_df[order(-LocalCC_ALLY_85),]
head(LocalCC_ALLY_85_df,10)
LocalCC_ALLY_95<- transitivity(g_ALLY_95, type="local")
LocalCC_ALLY_95_df<- data.frame(nodeName, LocalCC_ALLY_95)
LocalCC_ALLY_95_df <- LocalCC_ALLY_95_df[order(-LocalCC_ALLY_95),]
head(LocalCC_ALLY_95_df,10)
```
Countries with the same value of local clustering coefficient tends to be close to each other geographical and might in a same defense alliance network. (EG. Mideast, central america and north africa countries)

#Global clustering coefficient (CC)
global - ratio of triangles (direction disregarded) to connected triples.
```{r}
head(transitivity(g_ALLY_85),10)
head(transitivity(g_ALLY_95),10)
```

### Node-level Analysis

Considering the relationship between countries, we decided to look at the degree centrality(the communication between countries) and analyze betweenness centrality for ally network (which country has some unique position in forming alliances).

### Degree centrality

####1. War Network
This network represents coutries who have militarized dispute. 
```{r}
deg_WAR85 <- igraph::degree(g_WAR_85,mode="all")
V(g_WAR_85)$size <- deg_WAR85 * 0.5
E(g_WAR_85)$arrow.size <- 0.2
E(g_WAR_85)$edge$color <- "gray"
set.seed(1234)
plot(g_WAR_85,layout=layout.fruchterman.reingold)
```
```{r}
deg_WAR85 <- sort(deg_WAR85, decreasing = T)
head(deg_WAR85,10)
```

The USA most actively participates in war during 1985, then is Russia, Iran, Libya, Cuba, Uganda, Israel...
From the graph we could see more than half countries in our network are involved with wars, and most of them started a war with their neighbors or countries near them, and we think that is why Asian countries and African countries forming groups separately. Countries who are neutral, or didn`t participate in wars are spread around the group, like Czech or Austria as isolates. Those countries who have higher degree centrality are in the center of this network.


```{r}
deg_WAR95 <- igraph::degree(g_WAR_95,mode="all")
V(g_WAR_95)$size <- deg_WAR95 * 0.5
E(g_WAR_95)$arrow.size <- 0.2
E(g_WAR_95)$edge$color <- "gray"
set.seed(1234)
plot(g_WAR_95,layout=layout.fruchterman.reingold)
```

```{r}
deg_WAR95 <- sort(deg_WAR95, decreasing = T)
head(deg_WAR95,10)
```
Something worth notice here, in 1995, more countries have militarized dispute, which seems very different from 10 years ago. In 1995, Iran became the center of this war network, along with Iran and Turkey, they are surrounded by other countries. Another new center is Yugoslavia, who jumped to one of the countries who attended most of the militarized dispute. And the total number of the countries have ever attended in the war increases a lot. Which indicates the net is 'growing'.

#### 2. Defense Alliance Network

```{r}
deg_ALLY85 <- igraph::degree(g_ALLY_85,mode="all")
V(g_ALLY_85)$size <- deg_ALLY85 * 0.3
E(g_ALLY_85)$arrow.size <- 0.2
E(g_ALLY_85)$edge$color <- "gray"
set.seed(1234)
plot(g_ALLY_85,layout=layout.fruchterman.reingold)
```

```{r}
deg_ALLY85 <- sort(deg_ALLY85, decreasing = T)
deg_ALLY85
```
The graph shows 4 small groups who connected with each other, the USA and France located at the center of them, and they are surrounded by a lot of countries who are almost isolated.
The USA has most allies, then are Haiti, Dominican Republic, Jamacia, Trinidad and Tobago, Mexico, Guatemala, and some countries from North and Central America, we consider the countries who have 21 alliance had formed a united alliance, so they are allies with people who joined the alliance. The USA is in different alliances, so it has the highest allies number. Another interesting point is, some countries such as Japan and Australia, they only connect with the USA. Besides we also could see countries tend to make an alliance based on their geographical positions.

```{r}
deg_ALLY95 <- igraph::degree(g_ALLY_95,mode="all")
V(g_ALLY_95)$size <- deg_ALLY95 * 0.3
E(g_ALLY_95)$arrow.size <- 0.2
E(g_ALLY_95)$edge$color <- "gray"
set.seed(1234)
plot(g_ALLY_95,layout=layout.fruchterman.reingold)
```

```{r}
deg_ALLY95 <- sort(deg_ALLY95, decreasing = T)
deg_ALLY95
```
The USA has the most allies and France still served as an important communicative bridge between some country groups. In 1995, Canada appeared to make great connections with other countries, it became the second largest node in this network. Some ties emerging from country groups who used to be isolated like West and North Africa, maybe it`s because they are close to each other and try to make contact. Here France is the bridge linked Africa and Europe and America.

For most countries, after 10 years they grow more international connections, France and Canada are the most highly connected country in our network at 1995, something interesting is the ally of New Zealand and Kenya drops to 0 at 1995.

### Betweenness centrality for Ally Network

We think that in ally network, communication bridge is very important, so betweenness centrality is something worth discussing.

```{r}
bt_ALLY85 <- igraph::betweenness(g_ALLY_85)
V(g_ALLY_85)$size <- bt_ALLY85 * 0.01
E(g_ALLY_85)$arrow.size <- 0.2
E(g_ALLY_85)$edge$color <- "gray"
set.seed(1234)
plot(g_ALLY_85,layout=layout.fruchterman.reingold)
```

```{r}
bt_ALLY85 <- sort(bt_ALLY85, decreasing = T)
head(bt_ALLY85,10)
```

* Here we could clearly see who is the center in international defense connection. For America is USA, for Europe is France, for Africa is Djibouti and Senegal.


```{r}
bt_ALLY95 <- igraph::betweenness(g_ALLY_95)
V(g_ALLY_95)$size <- bt_ALLY95 * 0.01
E(g_ALLY_95)$arrow.size <- 0.2
E(g_ALLY_95)$edge$color <- "gray"
set.seed(1234)
plot(g_ALLY_95,layout=layout.fruchterman.reingold)
```

```{r}
bt_ALLY95 <- sort(bt_ALLY95, decreasing = T)
head(bt_ALLY95,20)
```
* After 10 years, more countries actively participated in forming alliances with other countries, Canada, Turkey, Azerbaijan poped up.

### Community detection

After compared with different algorithms, we choose to use the fast greedy algorithm to do our community detection analysis. Initially, every vertex belongs to a separate community, and communities are merged iteratively such that each merge is locally optimal. It is a way of modularity, the idea is to maximize the internal connection and better our clustering result.

```{r}
set.seed(1122)
fg <- fastgreedy.community(g_WAR_85)
plot (fg,g_WAR_85)
```

* Using fastgreedy algorithm, we detected 4 big groups and 4 small groups, those countries who didn`t participate in wars are isolated.

```{r}
set.seed(1122)
fg_WAR95 <- fastgreedy.community(g_WAR_95)
plot (fg_WAR95,g_WAR_95)
```

* Using fastgreedy algorithm, we detected 4 big groups and 5 small groups, they are more overlaped than countries in 1985, this is because more countries are involved but are not very active, still those countries who didn`t participate in wars are isolated.

```{r}
set.seed(1122)
fg_ALLY85 <- fastgreedy.community(g_ALLY_85)
plot (fg_ALLY85,g_ALLY_85)
```

* In 1985 ally network, we yield 5 groups, 1 is isolated, the other 4 are loosely connected with each other. Here we could see the countries have higher betweenness centrality served as 'bridges' between communities,


```{r}
set.seed(1122)
fg_ALLY95 <- fastgreedy.community(g_ALLY_95)
plot (fg_ALLY95,g_ALLY_95)
```

* In 1995 the network structure changed a little, we yield 5 groups who are loosely connected with each other. CHN and PRK participate into the ally communities. 'Bridges' connectors have also changed to countries who have higher betweenness centrality in their own community.


## Block modelling
### Data prep specifically for block modelling
```{r}
library(Matrix)
full_war_85 <- forceSymmetric(war_85,uplo="U")
full_war_85 <- as.matrix(full_war_85)
full_war_95 <- forceSymmetric(war_95,uplo="U")
full_war_95 <- as.matrix(full_war_95)
```

## Dendrogram for year 1985
```{r}
library(sna)
eq<-equiv.clust(list(full_war_85,ally_85), mode="digraph")
plot(eq)
```
  
Looking at the cluster dendrogram for 1985, we can see two interesting facts. First, the United States seems to be different from the other countries. This could stem from the fact that the US was very active geo-politically in both Latin America and Europe, whereas other countries are only geo-poltically active in regions close to them. Second, there are several countries that the dendgrogram does not split up at all. These are the countries that were neither involved in war, nor part of a defense alliance. Next, let us look at the block modelling image matrix.

## Create blockmodel for 1985
```{r}
b<-blockmodel(full_war_85,eq,k=4)

bimage <- b$block.model

fw85 <- network(full_war_85,directed=FALSE)
den <- network.density(fw85)

bimage[bimage < den] <- 0   
set.seed(1234)
gplot(bimage, diag=TRUE, 
      edge.lwd=bimage*0.03, 
      label=colnames(bimage),
      vertex.cex=sqrt(table(b$block.membership))/2,
      gmode="digraph", vertex.sides=50,label.pos = 4,
      vertex.col=gray(1-diag(bimage)/2))
```
  
In the image matrix above, we see four blocks. Block 1 is connected to itself and to Block 2. Block 2 is connected to itself and Block 1 and 4. Block 4 is connected to itself and Block 3 and 2. Block 3 is only connected to itself. To get a better idea of how to interpret this, let us look at which countries are in which block.

## Mapping countries to blocks in 1985
```{r}
# Get the countries from block 1
group1 <- b$order.vector[b$block.membership == 1]
colnames(full_war_85)[group1]

# Get the countries from block 2
group2 <- b$order.vector[b$block.membership == 2]
colnames(full_war_85)[group2]

# Get the countries from block 4
group4 <- b$order.vector[b$block.membership == 4]
colnames(full_war_85)[group4]

# Get the countries from block 3
group3 <- b$order.vector[b$block.membership == 3]
colnames(full_war_85)[group3]

# Plot image matrix with labels
set.seed(1234)
gplot(bimage, diag=TRUE, 
      edge.lwd=0.3, 
      label=c('Block 1 - Americas', 'Block 2 - \nWestern Europe',  'Block 3 - \nRest of the world', 'Block 4 - \nNorth Africa/Middle East'),
      vertex.cex=sqrt(table(b$block.membership))/2,
      gmode="digraph", vertex.sides=50, label.pos = 4,
      vertex.col=gray(1-diag(bimage)/2))
```
  
Once we include the labels, the interpretation becomes much easier. The Americas are connected to each other because of several wars and defense alliances going on in 1985. The Americas are also connected to Western Europe, mostly through the ties of the United States. Western Europe is connected to itself through defense alliances and to North Africa/Middle East because Western European countries had many colonies in North Africa/Middle East. North Africa/Middle East is connected to itself because of several wars going on in the region at the time. Overall, the block model highlights several interesting aspects of geo-political activities in 1985.

## Dendrogram for year 1995
```{r}
eq<-equiv.clust(list(full_war_95,ally_95), mode="digraph")
plot(eq)
```
  
The dendrogram of 1995 shows one significant difference compared to 1985. The United States are not split up separately at a high level anymore. This shows that the United States reduced its geo-political activities between 1985 and 1995. Let us see how that influenced the block modelling image matrix.

## Create block model for 1995 and map countries to blocks
```{r}
b<-blockmodel(full_war_95,eq,k=4)

bimage <- b$block.model

fw95 <- network(full_war_95,directed=FALSE)
den <- network.density(fw95)

bimage[bimage < den] <- 0   

# Get the countries from block 1
group1 <- b$order.vector[b$block.membership == 1]
colnames(full_war_95)[group1]

# Get the countries from block 2
group2 <- b$order.vector[b$block.membership == 2]
colnames(full_war_95)[group2]

# Get the countries from block 3
group3 <- b$order.vector[b$block.membership == 3]
colnames(full_war_95)[group3]

# Get the countries from block 4
group4 <- b$order.vector[b$block.membership == 4]
colnames(full_war_95)[group4]

# Plot image matrix with labels
set.seed(1001)
gplot(bimage, diag=TRUE, 
      edge.lwd=bimage*0.03, 
      label=c('Block 1 - Americas', 'Block 2 - Rest of the world',  'Block 3 - Europe', 'Block 4 - North Africa/Middle East'),
      vertex.cex=sqrt(table(b$block.membership))/2,
      gmode="digraph", vertex.sides=50, label.pos = 2,
      vertex.col=gray(1-diag(bimage)/2))
```
  
Compared to the image matrix for 1985, we can see several significant changes. First, the Americas are not connected to Europe anymore. As we saw in the dendrogram for 1995, this is because the United States stopped being geo-politically active in Europe. Second, the North Africa/Middle East is not connected to the rest of the world anymore. Western Europe is now connected to the rest of the world though and is also still connected to North Africa/Middle East. As mentioned before, this is because Europe had many former colonies in North Africa/Middle East.

# TERGM model

Unlike the previous sections, we used the data for the entire ten-year period between 1985 and 1995 to give our model better information. Here, we create subset of all the networks for the mentioned time period. 

```{r}
allyNet_85_95 <- allyNet[c(5:15)]

#converting to networkdynamic data
allydyn_85_95 <- networkDynamic(network.list = allyNet[c(5:15)])
#we use the war data for 1984-1994 because we want information the war in the previous year
warNet_84_94 <- warNet[c(4:14)]
lNet_85_95 <- lNet[c(5:15)]
for (i in 1:11){
  warNet_84_94[lower.tri(warNet_84_94)] = t(warNet_84_94)[lower.tri(warNet_84_94)]
  lNet_85_95[lower.tri(lNet_85_95)] = t(lNet_85_95 )[lower.tri(lNet_85_95)]
}

```

We use the btergm package since this creates bootstrapped samples and creates the tergm model based on these repeated samples. Based on repeated iterations, we found the following model to have the least error. We ran the bootstrapping taking 50 samples from the alliance network between the years 1985-1995.


```{r}
 allymodel <- btergm(allyNet_85_95 ~ edges + triangles + gwesp(0, fixed = TRUE) + nodecov("polity") +
+ nodecov("cinc") + absdiff("polity") + absdiff("cinc") +
+ edgecov(contigMat)  
+ edgecov(lNet_85_95)
+ edgecov(warNet_84_94), R = 50)
```

```{r}
summary(allymodel)
```
Since we are not given a p-value in this summary, we estimate statistical significance by seeing if 0 is contained within the 95% confidence interval of the coefficient estimates. Although this is not the ideal method to estimate statistical significance, it will suffice for the purposes of this project. 

Based on the above method, we find that nodecov.polity and absdiff.cinc are not significant. absdiff.polity is nearly 0.

### Let us discuss the interpretations of the significant predictors :
* We see that edge has a very negative beta coefficient. This makes sense as our network is sparse and so our mean density will also be less. 
* We notice that triangles has a positive coefficent. This tells us that a change in the number of triangles involving two nodes iuncreases the probability of those two nodes forming an alliance.
* We see that the geometrically weighted edgewise shared partners term is a strong predictor of forming ties. This also makes sense as we know that sharing an ally will increase the chances of two countries sharing an ally, but this increase in probability will decay and hence we use a geometrically weighted term instead of just "edge-wise shared partners".
* The edge covariance of the LSP matrix tells us the number of shared partner between countries in the **previous year** and thus it gives us slightly different information from the gwesp term. The fact that even this term has a high coefficent tells us that both current shraed partners and a history of sharing partners is important in forming ties. 
* CINC gives us an indication of state capabilities. It makes sense that this is the strongest indicator of forming a tie. A state that is more capable will be more in demand and hence will have more allies. Other less capable countries may want to align themselves with a strong states to seek their protection and help. Especially during the 1985-1995 period which involved the splitting up of USSR, each of the smaller units would want have wanted to align themselves with a strong power. We notice that absdiff.cinc was not significant which means that both capable and less capable country. would want to ally with a more capable country, 
* We see that countries that share a border are extremely likely to be allies. This is intuitive as geographic neighbours would seek out each other and collectively defend their borders against outside intrusion. 
* We see that countries that were allied in the previous year have a very high probabolity of being allie din the current year. Like most time series data, the alliance network is also highly autocorrelated. 
* Two countries that were at war in the previous year are less likely to form an alliance in the current year. This also seems intuitive. However, we would have expected this factor to have a higher value. This could be because some countries may enter into an alliance after a dispute has ended as a part of a treaty of some sort.

### Goodness of Fit Tests

We test the goodness of fit of our model by plotting the degree of simulated networks 

```{r}
gofmodel <- gof(allymodel, nsim = 50, statistics = c(deg))
```

```{r}
plot(gofmodel)
```

We see that our model does a decent job of simulating degree of alliances in the network since the thick black line showing the observed network degree distribution is close to the medians in the box plots of the simulated networks. 

## Discussion and Conclusion 

From the above discussion, we found some interesting patterns in what drove wars and alliances 
in 1985-1995. These patterns consistently showed up in all the types of analysis we performed. 

* Despite 1985-1995 being a highly untable period in world history, the edge density in both war and alliance networks is not too high which tells us that a large portion of the countries in the world were at peace. 
* Since the period across 1985-1995 involved a lot of political upheaval, we see a lot of changes in the alliance and war network, with shifting importances and communities. 
* Geography and sharing borders is a huge factor in both wars and alliances. Countries tend to have disputes with their neighbours but when an external force threatens them, they also tend to seek their neighbour's support to collectively protect their borders. 
* Like most real networks, there are a few countires like USA snd France that are more important and have higher degrees than other countires. These countries also link to the smaller countries un the Americas (in the case of USA) and Africe (in France's case) and form subgroups of alliances and wars in those regions. 
* "Important" countires, that is, those with the highest centralities seem to be the countries with highest CINC scores. Thus the "most capable"" states are also attractive allies and also the most active in the geopolitical field. 
* Most of war and alliances is about forming "teams" or "sides". Thus, sharing partners is extremely important for forming alliances. Countries form alliances with those countries with whom they share alliances. 





